---
title: "Validation checks"
author: "Graham French"
date-modified: "`r lubridate::today()`"
format: html
execute: 
  eval: false
knitr:
  opts_chunk:
    comment: "#>"
editor: visual
---

```{r setup}
suppressPackageStartupMessages({
  suppressWarnings({
    library(dplyr)
    library(assertr)
    library(assertive.properties)
    library(assertive.strings)
    library(assertive.types)
    library(assertthat)
    library(palmerpenguins)
    library(data.validator)
  })
})
```

## Validation checks

Data validation is performed using the [assertr](https://cran.r-project.org/web/packages/assertr/vignettes/assertr.html) package, with extra checks provided by the assertive package.

::: {.callout-warning appearance="simple"}
The assertive package and other associated assertive package was archived on CRAN on 2024-10-23.

To use the following functions the associated assertive package will need to be installed from the CRAN archive in the following order

```{r}
#| eval: FALSE

## is_true
## is_not_na
remotes::install_version(package = "assertive.base", version = "0.0-9", repos = "https://cloud.r-project.org")

## has_elements
remotes::install_version(package = "assertive.properties", version = "0.0-5", repos = "https://cloud.r-project.org")

## is_an_integer
remotes::install_version(package = "assertive.types", version = "0.0-3", repos = "https://cloud.r-project.org")

## is_matching
## is_matching_regex
## is_not_matching_regex
remotes::install_version(package = "assertive.strings", version = "0.0-3", repos = "https://cloud.r-project.org")

## is_before
## is_after
remotes::install_version(package = "assertive.datetimes", version = "0.0-3", repos = "https://cloud.r-project.org")

```
:::

Appsilon's [data.validator](https://appsilon.github.io/data.validator/index.html) package is used to create a validation report or export the validation results in a spreadsheet.

The [assertthat](https://github.com/hadley/assertthat) package can also be used as an alternative to stopifnot().

## Directory

```{r}
assert_that(file.exists("validation_checks.qmd"), 
            msg = "Path to validation check document is UNSUCCESSFUL")
```

## Dataset

```{r}
adelie <- penguins |> 
  filter(species == "Adelie", !is.na(bill_length_mm)) |> 
  mutate(id = row_number(), .before = everything()) |> 
  mutate(across(where(is.factor), as.character))
```

```{r}
see_if(not_empty(adelie), 
       msg = "Adelie dataset is EMPTY")
```

```{r}
report <- data_validation_report()
```

```{r}
adelie |> 
  validate(name = "Dataset") |> 
  validate_if(description = "Check dataset contains a specific number of rows",
              expr = has_elements(id, 151)) |> 
  validate_if(description = "Check for column names in dataset",
              expr = has_only_names(c("id", "species", "island", "bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g", "sex", "year"))) |>
  validate_cols(description = "Check for data type of one or more columns",
              predicate = is_an_integer,
              body_mass_g) |> 
  add_results(report)
```

## Values

### Unique values

```{r}
adelie |> 
  validate(name = "Unique values") |>
  validate_if(description = "Check for unique values in a single column",
              expr = is_uniq(id)) |>
  validate_rows(description = "Check for unique values across multiple columns",
                row_reduction_fn = col_concat,
                predicate = is_uniq,
                everything()) |>
  add_results(report)
```

### Missing values

```{r}
#| column: page-inset-right

adelie |> 
  validate(name = "Missing values") |>
  validate_if(description = "Check for missing values in a single column",
              expr = not_na(species)) |>
  validate_if(description = "Check for non missing values in a single column",
              expr = is.na(species)) |>
  validate_cols(description = "Check for any missing values in multiple columns",
                predicate = not_na,
                species, island) |> 
  validate_rows(description = "Check for all missing values in multiple columns",
                row_reduction_fn = num_row_NAs, # function (x) {sum(is.na(x)}
                predicate = within_bounds(lower.bound = 0, upper.bound = 4, include.upper = FALSE),
                bill_length_mm:body_mass_g) |> 
  validate_cols(description = "Check for non missing values in multiple columns",
                predicate = is.na,
                species, island) |> 
  validate_rows(description = "Check for number of missing values across multiple columns",
                row_reduction_fn = num_row_NAs,
                predicate = in_set(0),
                bill_length_mm:body_mass_g) |>
  add_results(report)
```

### Restricted values

```{r}
#| column: page-inset-right

adelie |> 
  validate(name = "Restricted values") |> 
  validate_if(description = "Warn if values in a single column are above a defined value or are missing",
              skip_chain_opts = TRUE, error_fun = warning_append,
              expr = bill_length_mm < 50) |> 
  validate_cols(description = "Warn if values in one or more columns are outside a given range",
                skip_chain_opts = TRUE, error_fun = warning_append,
                predicate = within_bounds(lower.bound = 30, upper.bound = 50),
                bill_length_mm) |>
  validate_if(description = "Check for compulsory value in a single column",
              expr = is_matching_fixed(species, "Adelie")) |> # is_matching_regex(PATTERN)
  validate_if(description = "Check for compulsory substring value a single column",
              expr = is_matching_regex(sex, "male") | is.na(sex)) |>
  validate_cols(description = "Check for compulsory restricted values in one or more columns (supply values as an external vector for four or more values; is_matching_fixed | is_matching_regex does not apply as a predicate in validate_cols for across more than one column)",
                predicate = in_set("male", "female", allow.na = TRUE),
                sex) |>
  validate_cols(description = "Check for compulsory restricted values not in one or more columns",
                predicate = in_set("TBC", allow.na = FALSE, inverse = TRUE),
                sex) |>
  add_results(report)
```

## Validation report

```{r}
#| eval: FALSE
unnest_failed_validation_results(report)

# save report results in error and has been commented out
# save_report(report)
# save_report(report) # remove Registered S3 method message at top of report
# browseURL("validation_report.html")
```
