---
title: "Create and update a package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create and update a package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval=FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r packages, echo=FALSE}
library(store)
```

This documents the functions to run to help create and update a R package. Further information can be found at

* [devtools documentation](https://devtools.r-lib.org/)
* [usethis documentation](https://usethis.r-lib.org/)
* [pkgdown documentation](https://pkgdown.r-lib.org/)
* [renv documentation](https://rstudio.github.io/renv/articles/renv.html)

## Create the skeleton package --------------------------------------------

**usethis::create_package** create the skeleton package directory structure, opening the package in a new R Studio session as a new project
```{r}
usethis::create_package("C:/Files/packages/store")
```

**usethis::use_description** update the title, author and description details in the description file
```{r}
usethis::use_description(fields = list(Title = "Store commonly used functions and lookups",
                                       `Authors@R` = person("Graham", "French", 
                                                            role = c("aut", "cre"),
                                                            email = "gcfrench@googlemail.com"),
                                       Description = "This package stores the commonly used
                                       functions and lookups, along with function tests and
                                       documentation. These can be then used when creating new
                                       packages by copy and pasting the relevant R scripts and
                                       associated test scripts into the new package, as well as
                                       updating the pkgdown file",
                                       URL = "https://gcfrench.github.io/store"))
```

**usethis::use_cc0_license** add one of the license options for the package
```{r}
usethis::use_cc0_license()
```

**usethis::use_readme_rmd** add a readme file, removing the default text from the file. This file is displayed as the front home page in the pkgdown website
```{r}
usethis::use_readme_rmd()
```

**usethis::use_news_md** add a news file, removing the default text from the file. This file will contain the news on each version release
```{r}
usethis::use_news_md()
```

**usethis::use_pipe** add the %>% pipe operator with the addition of the utils-pipe.R file and magrittr package to the imports field in the description file
```{r}
usethis::use_pipe()
```

**usethis::use_data_raw** add the data-raw directory and empty lookup_raw containing the scripts used to format and add lookup datasets
```{r}
usethis::use_data_raw("lookup_raw")
```

**usethis::use_testthat** add the testthat directory used to contain the test scripts and testthat package to suggests field in the description file
```{r}
usethis::use_testthat()
```

**devtools::document** add the man directory containing the document .Rd files. Documents can also be updated using Ctlr-shift-D or through the Build:Document menu
```{r}
devtools::document()
```

**usethis::use_pgkdown** add the _pkgdown.yml file used to create the menu structure for the package website
```{r}
usethis::use_pkgdown()
```

pkgdown.yml template
```{}
template:
  params:
    bootswatch: flatly
    
reference:
- title: ""
  desc: ""
- contents:
  - function_name
  
navbar:
  title: package_name
  left:
    - icon: fa-home fa-lg
      href: index.html
    - text: "information"
      menu:
      - text: ""
        href: articles/script_name.html
    - text: "lookups and functions"
      href: reference/script_name.html
    - text: "analyses"
      menu:
      - text: ""
        href: articles/script_name.html
    - text: "versions"
      href: news/index.html
   right:
    - icon: "fab fa-twitter fa-lg"
      href: https://twitter.com/gcfrench
    - icon: "fab fa-github fa-lg"
      href: https://github.com/gcfrench/store
```

**usethis::use_lifecycle** add the lifecycle badges in the man directory
```{r}
usethis::use_lifecycle()
```

**usethis::use_lifecycle_badge** add lifecycle badge to home page
```{r}
usethis::use_lifecycle_badge("experimental") # "experimental", "stable". "superseded", "deprecated"
```

**usethis::use_code_of_conduct** add code of conduct markdown, pasting code of conduct section into README file
```{r}
usethis::use_code_of_conduct()
```

## Add to github --------------------------------------------

**usethis::use_git** initiates a Git repository, making the initial commits
```{r}
usethis::use_git()
```

**usethis::use_github** creates the repository on Github and pushes the initial commits. This needs a Github Personal Access Token (PAT) which can be created using **usethis::create_github_token**. Information on how to do this can be found in [Happy Git and Github for the useR](https://happygitwithr.com/credential-caching.html)
```{r}
usethis::use_github()
# usethis::create_github_token()
```

**usethis::use_github_action** configures files to use github to host package website
```{r}
usethis::use_github_action("pkgdown")
```

## Add package versioning --------------------------------------------

Navigate to the home project directory first

**renv::init** initiates the renv environment, creating the renv folder and adding the renv.lock file, linking the relevant packages from the renv cache
```{r}
renv::init()
```

**renv::status** check the renv lock file is up-to-date, run **renv::snapshot** to update the renv lock file if not

```{r}
renv::status()
renv::snapshot()
```

**renv::deactivate** turn off the renv environment, then manually delete the renv folder and .Rprofile file
```{r}
renv::deactivate()
```

Navigate to the home project directory first

**renv::activate** turn on the renv environment
```{r}
renv::activate()
```

**renv::restore** restore the package versions contained within the renv lock file
```{r}
renv::restore()
```

## Add and edit script files

### Add new script --------------------------------------------

**usethis::use_vignette**("vignette_name"): adds the vignette directory, containing a new vignette. This directory contains the script files. Add eval=False to the knitr::opts_chunk to turn off running the code.
```{r}
usethis::use_vignette("script_name")
```

**pkgdown::build_articles** builds the html files in the articles sub-directory of the docs directory. Update the _pkgdown.yml to contain the link to the html file
```{r}
pkgdown::build_article("script_name")
# pkgdown::build_articles()
pkgdown::build_articles_index()
```

### Add new function --------------------------------------------

**usethis::use_r** add new r script in the R directory containing functions. Run and edit Roxygen comments using Code:Insert Roxygen skeleton, or Ctlr-Alt-Shift-R. update documentation by runing **devtools::document**, Ctlr-shift-D or through the Build:Document menu
```{r}
usethis::use_r("function_script_name")
```

**usethis::use_package** add packages to the imports section of the description file
```{r}
usethis::use_package("package_name")
```

**devtools::document** update documentation, can also use Ctlr-shift-D or through the Build:Document menu
```{r}
devtools::document()
```

**pkgdown::build_reference** builds the html files in the reference sub-directory of the docs directory. Update the _pkgdown.yml to contain the link to the html file
```{r}
pkgdown::build_reference
pkgdown::build_reference_index()
```

### Add new test script --------------------------------------------

**usethis::use_test** add unit tests for function
```{r}
usethis::use_test("test_script_name")
```

**devtools::test** run all tests, can also use Ctrl-Shift_T or Build-Test Package
```{r}
devtools::test()
```

### Add new data script --------------------------------------------

**usethis::use_data_raw** add a new data_raw script
```{r}
usethis::use_data_raw("data_raw_script_name")
```

**usethis::use_data** add the .rda data file in the data directory, after running the data raw script so that the data frame is stored in the global environment
```{r}
usethis::use_data(data_frame_name)
```

**usethis::use_r** document the data frame in a R script. Running **devtools::document** afterwards
```{r}
usethis::use_r("lookup")
```

Data frame documentation template
```
#' Data frame name
#'
#' Data frame description
#'
#' @format A tibble with xx rows and x variables
#' \describe{
#'   \item{field_name}{field description}
#'   \item{field_name}{field description}
#' }
"data_frame_name"
```

**pkgdown::build_reference** builds the html files in the reference sub-directory of the docs directory. Update the _pkgdown.yml to contain the link to the html file
```{r}
pkgdown::build_reference
pkgdown::build_reference_index()
```

### Edit news file --------------------------------------------

**usethis::use_version** update the package version
```{r}
usethis::use_version()
```

**usethis::edit_file** edit the news file with the new package version
```{r}
usethis::edit_file("News.md")
```

**pkgdown::build_news** builds the index.html file in the news sub-directory of the docs directory
```{r}
pkgdown::build_news()
```

### Edit the README file --------------------------------------------

**usethis::edit_file** edit the README.Rmd file
```{r}
usethis::edit_file("README.Rmd")
```

**pkgdown::build_home** builds the index.html file in the docs directory
```{r}
pkgdown::build_home()
```

## Publish on GitHub

**pkgdown::deploy_to_branch** pushes docs folder to GitHub branch gh-pages. 

Need to change repository settings GitHub Pages Source to Branch gh-pages/root. 
Site published to **https://github_name.github.io/package_name**
```{r}
pkgdown::deploy_to_branch()
```

**usethis::edit_file** Add URL to description and README files
```{r}
usethis::edit_file("DESCRIPTION")
usethis::edit_file("README.Rmd")
```

## Install package

Build documentation, test functions and build website, then run **Install and Restart** to build and install package

```{r}
devtools::document()
devtools::test()
pkgdown::build_site()
```





