---
title: "Create and update a shiny package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create and update a shiny package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

This documents the functions to run to help create and update a R shiny package. Further information can be found at

-   [golem documentation](https://thinkr-open.github.io/golem/)

## Create the skeleton package

**golem::create_golem** create the skeleton shiny package directory structure, opening the package in a new R Studio session as a new project.

```{r}
golem::create_golem("C:/Files/packages/shinystore")
```

**golem::fill_desc** update the title, author and description details in the description file
```{r}
golem::fill_desc(
  pkg_name = "shinystore", # The Name of the package containing the App
  pkg_title = "Stores a shiny example app using modules", # The Title of the package containing the App
  pkg_description = "This shiny package is an example shiny app created using the 
  golem package and using shiny modules. It demonstrates many of the shiny techniques 
  described in Hadley Wickham's Mastering Shiny book", # The Description of the package containing the App
  author_first_name = "Graham", # Your First Name
  author_last_name = "French", # Your Last Name
  author_email = "gcfrench@googlemail.com", # Your Email
  repo_url = "https://gcfrench.github.io/shinystore" # The URL of the GitHub Repo (optional)
)
```

**usethis::use_cc0_license** add one of the license options for the package
```{r}
usethis::use_cc0_license()
```

**usethis::use_readme_rmd** add a readme file, removing the default text from the file. This file is displayed as the front home page in the pkgdown website
```{r}
usethis::use_readme_rmd()
```

```{}
## Overview

## Installation

## Usage

## Publication
```

