---
title: "Create and update a package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Create and update a package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval=FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r packages, echo=FALSE}
library(store)
```

This documents the functions to run to help create and update a R package. Further information can be found at

* [devtools documentation](https://devtools.r-lib.org/)
* [usethis documentation](https://usethis.r-lib.org/)
* [pkgdown documentation](https://pkgdown.r-lib.org/)
* [renv documentation](https://rstudio.github.io/renv/articles/renv.html)
* [golem documentation](https://thinkr-open.github.io/golem/)

## Set environment ---------------

**usethis::use_blank_slate** turn off saving and restoring workspace between sessions

```{r}
usethis::use_blank_slate(scope = "project")
```

## Create the skeleton package ---------------

**usethis::create_package** or **golem::create_golem** create the skeleton package directory structure, opening the package in a new R Studio session as a new project
```{r}
usethis::create_package("C:/Files/packages/store")
golem::create_golem("C:/Files/packages/shinystore")
```


**usethis::use_description** or **golem::fill_desc** update the title, author and description details in the description file
```{r}
usethis::use_description(fields = list(Title = "Store commonly used functions and lookups",
                                       `Authors@R` = person("Graham", "French", 
                                                            role = c("aut", "cre"),
                                                            email = "gcfrench@googlemail.com"),
                                       Description = "This package stores the commonly used
                                       functions and lookups, along with function tests and
                                       documentation. These can be then used when creating new
                                       packages by copy and pasting the relevant R scripts and
                                       associated test scripts into the new package, as well as
                                       updating the pkgdown file",
                                       URL = "https://gcfrench.github.io/store"))

golem::fill_desc(
  pkg_name = "shinystore", # The Name of the package containing the App
  pkg_title = "Stores a shiny example app using modules", # The Title of the package containing the App
  pkg_description = "This shiny package is an example shiny app created using the 
  golem package and using shiny modules. It demonstrates many of the shiny techniques 
  described in Hadley Wickham's Mastering Shiny book", # The Description of the package containing the App
  author_first_name = "Graham", # Your First Name
  author_last_name = "French", # Your Last Name
  author_email = "gcfrench@googlemail.com", # Your Email
  repo_url = "https://gcfrench.github.io/shinystore" # The URL of the GitHub Repo (optional)
)
```

**usethis::use_cc0_license** add one of the license options for the package
```{r}
usethis::use_cc0_license()
```

**usethis::use_readme_rmd** add a readme file, removing the default text from the file. This file is displayed as the front home page in the pkgdown website
```{r}
usethis::use_readme_rmd()
```

```{}
## Overview

## Installation

## Usage

## Publication
```

**usethis::use_news_md** add a news file, removing the default text from the file. This file will contain the news on each version release
```{r}
usethis::use_news_md()
```

```{}
## package_name 0.0.0.9000 <font size="4">yyyy-mm-dd</font>
```

**usethis::use_pipe** add the %>% pipe operator with the addition of the utils-pipe.R file and magrittr package to the imports field in the description file
```{r}
usethis::use_pipe()
```

**usethis::use_data_raw** add the data-raw directory and empty lookup_raw containing the scripts used to format and add lookup datasets
```{r}
usethis::use_data_raw("lookup_raw")
```

**usethis::use_testthat** add the testthat directory used to contain the test scripts and testthat package to suggests field in the description file
```{r}
usethis::use_testthat()
```

**golem::use_recommended_tests** add the test folder and golem tests
```{r}
golem::use_recommended_tests()
```

**usethis::use_coverage** or **covr::package_coverage** - adds test coverage using 
either Codecov or Coveralls webservice
```{r}
usethis::use_coverage(type = "codecov")
usethis::use_coverage(type = "coveralls")
```

**devtools::document** add the man directory containing the document .Rd files. Documents can also be updated using Ctlr-shift-D or through the Build:Document menu
```{r}
devtools::document()
```

**usethis::use_vignette** add tasks vignette
```{r}
usethis::use_vignette("tasks")
```

```{}
This **tasks** section contains a list of tasks and questions, with
their current status

-   ✔️ completed
-   ❗ awaiting completion
-   ✖️no longer needed
-   ❓question to be answered

## Tasks and questions

### YYYY-MM-DD
```

**usethis::use_pkgdown** add the _pkgdown.yml file used to create the menu structure for the package website
```{r}
usethis::use_pkgdown()
pkgdown::init_site()
```

**pkgdown::build_article** create tasks html report
```{r}
pkgdown::build_article("tasks")
```

**create yaml template** pkgdown.yml template
```{}
title: TITLE_HERE
template:
  bootstrap: 5
  bootswatch: flatly

navbar:
  left:
    - icon: fa-home fa-lg
      href: index.html
    - text: information
      menu:
      - text:
        href:
    - text: "lookups and functions"
      href: reference/index.html
    - text: "archive"
      menu:
      - text:
        href:
    - text: "tasks"
      href: articles/tasks.html
    - text: "versions"
      href: news/index.html
  right:
    - icon: "fab fa-github fa-lg"
      href: https://github.com/URL_HERE
```

**usethis::use_roxygen_md** uses roxygen2 with markdown
```{r}
usethis::use_roxygen_md()
```

**usethis::use_lifecycle** add the lifecycle badges in the man directory
```{r}
usethis::use_lifecycle()
```

**usethis::use_lifecycle_badge** add lifecycle badge to home page
```{r}
usethis::use_lifecycle_badge("experimental") # "experimental", "stable". "superseded", "deprecated"
```

**usethis::use_logo** or **golem::use_favicon** add a hex logo to home page
```{r}
# create hex logo using hexmake https://connect.thinkr.fr/hexmake
usethis::use_logo("image_path")

# add favicon
golem::use_favicon( path = "path/to/favicon")
# golem::remove_favicon()
```

**RStudio: Build - Install and Restart**

**pkgdown::build_site** build documentation
```{r}
pkgdown::build_site()
```

### Add to github ---------------

**usethis::use_git** initiates a Git repository, making the initial commits
```{r}
usethis::use_git()
```

**usethis::use_github** creates the repository on Github and pushes the initial commits. This needs a Github Personal Access Token (PAT) which can be created using **usethis::create_github_token**. Information on how to do this can be found in [Happy Git and Github for the useR](https://happygitwithr.com/credential-caching.html)
```{r}
usethis::use_github()
# usethis::create_github_token()
```

**usethis::use_github_action** configures files to use github to host package website
```{r}
usethis::use_github_action("pkgdown")
```

### Add config file ---------------

**usethis::edit_file** add a config file and package to descriptions
```{r}
usethis::edit_file("config.yml")
usethis::use_package("config")
```

```{}
default:

testing:

production:
```

**golem::set_golem_options** update golem-config file with description information
and set here path
```{r}
golem::set_golem_options()
```

### Add package versioning ---------------

Navigate to the home project directory first

**renv::init** initiates the renv environment, creating the renv folder and adding the renv.lock file, linking the relevant packages from the renv cache
```{r}
renv::init()
```

**renv::status** check the renv lock file is up-to-date, run **renv::snapshot** to update the renv lock file if not

```{r}
renv::status()
renv::snapshot()
```

**renv::deactivate** turn off the renv environment, then manually delete the renv folder and .Rprofile file
```{r}
renv::deactivate()
```

Navigate to the home project directory first

**renv::activate** turn on the renv environment
```{r}
renv::activate()
```

**renv::restore** restore the package versions contained within the renv lock file
```{r}
renv::restore()
```

## Add and edit script files

### Add documentation folders ---------------
```{r}
fs::dir_create("man/figures")
fs::dir_create("man/examples")
```

### Add new script ---------------

**usethis::use_vignette**("vignette_name"): adds the vignette directory, containing a new vignette. This directory contains the script files. Add eval=False to the knitr::opts_chunk to turn off running the code.
```{r}
usethis::use_vignette("script_name")
```

**pkgdown::build_articles** builds the html files in the articles sub-directory of the docs directory. Update the _pkgdown.yml to contain the link to the html file
```{r}
pkgdown::build_article("script_name")
# pkgdown::build_articles()
pkgdown::build_articles_index()
```

### Add new function ---------------

**usethis::use_r** or **golem::add_fct** or **golem::add_utils**  add new r script in the R directory containing functions. Run and edit Roxygen comments using Code:Insert Roxygen skeleton, or Ctlr-Alt-Shift-R. update documentation by runing **devtools::document**, Ctlr-shift-D or through the Build:Document menu
```{r}
usethis::use_r("function_script_name")

# files containing small functions used several times in the application
golem::add_utils("function_script_name")

# files containing larger functions that are more central to the application
golem::add_fct("function_script_name") 
```

**usethis::use_package** add packages to the imports section of the description file
```{r}
usethis::use_package("package_name")
```

**golem::use_recommended_deps** add recommended package dependencies
```{r}
golem::use_recommended_deps()
```

**devtools::document** update documentation, can also use Ctlr-shift-D or through the Build:Document menu
```{r}
devtools::document()
```

**pkgdown::build_reference** builds the html files in the reference sub-directory of the docs directory. Update the _pkgdown.yml to contain the link to the html file
```{r}
pkgdown::build_reference
pkgdown::build_reference_index()
```

### Add utility functions ---------------

**golem::use_utils_ui** and **golem::use_utils_server** adds golem utility functions
```{r}
golem::use_utils_ui()
golem::use_utils_server()
```

### Add new module ---------------

**golem::add_module** add module in R folder
```{r}
golem::add_module(name = "MODULE_NAME")
```

### Add new test script ---------------

**usethis::use_test** add unit tests for function
```{r}
usethis::use_test("test_script_name")
```

**devtools::test** run all tests, can also use Ctrl-Shift_T or Build-Test Package
```{r}
devtools::test()
```

**covr::package_coverage** - returns test coverage of R scripts
```{r}
covr::package_coverage()
```

### Add new data script ---------------

**usethis::use_data_raw** add a new data_raw script
```{r}
usethis::use_data_raw("data_raw_script_name")
```

**usethis::use_data** add the .rda data file in the data directory, after running the data raw script so that the data frame is stored in the global environment
```{r}
usethis::use_data(data_frame_name)
```

**usethis::use_r** document the data frame in a R script. Running **devtools::document** afterwards
```{r}
usethis::use_r("lookup")
```

Data frame documentation template
```
#' Data frame name
#'
#' Data frame description
#'
#' @format A tibble with xx rows and x variables
#' \describe{
#'   \item{field_name}{field description}
#'   \item{field_name}{field description}
#' }
"data_frame_name"
```

**pkgdown::build_reference** builds the html files in the reference sub-directory of the docs directory. Update the _pkgdown.yml to contain the link to the html file
```{r}
pkgdown::build_reference
pkgdown::build_reference_index()
```

### Add new external files

**golem::add_js_file** add javascript file to inst/app/www folder
```{r}
golem::add_js_file("script")
golem::add_js_handler("handlers")
```

**golem::add_css_file** add blank css file to inst/app/www folder
```{r}
golem::add_css_file("custom")
```

### Edit news file ---------------

**usethis::use_version** or **golem::set_golem_version** update the package version
```{r}
usethis::use_version()
golem::set_golem_version()
```

**usethis::edit_file** edit the news file with the new package version
```{r}
usethis::edit_file("News.md")
```

**pkgdown::build_news** builds the index.html file in the news sub-directory of the docs directory
```{r}
pkgdown::build_news()
```

### Edit the README file ---------------

**usethis::edit_file** edit the README.Rmd file
```{r}
usethis::edit_file("README.Rmd")
```

**pkgdown::build_home** builds the index.html file in the docs directory
```{r}
pkgdown::build_home()
```

## Install package ---------------

Build documentation and test functions, then run either **Install and Restart** 
or **remotes::install_local** to build and install package

```{r}
devtools::document()
devtools::test()
remotes::install_local()
```

## Publish on GitHub ---------------

**pkgdown::build_site** run **pkgdown::clean_site** first to delete files in docs
sub-directory
```{r}
pkgdown::clean_site()
pkgdown::build_site()
```

**pkgdown::deploy_to_branch** pushes docs folder to GitHub branch gh-pages. 

Need to change repository settings GitHub Pages Source to Branch gh-pages/root. 
Site published to **https://github_name.github.io/package_name**
```{r}
pkgdown::deploy_to_branch()
```

**usethis::edit_file** Add URL to description and README files
```{r}
usethis::edit_file("DESCRIPTION")
usethis::edit_file("README.Rmd")
```

## Send to another machine ---------------


**pkgbuild::build** create a tar.gz file to send to other machine
```{r}
pkgbuild::build()
remotes::install_local("path/to/tar.gz")
```

## Deploy on shinyapp.io ---------------

**golem::add_shinyappsio_file** adds an app.R file
```{r}
golem::add_shinyappsio_file()
```

## Create a docker file ---------------

**golem::add_dockerfile** creates a Dockerfile
```{r}
golem::add_dockerfile()
```






