---
title: "Git notes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Git notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval=FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

These notes were made whilst going through Jenny Bryan's [Happy Git and GitHub for the useR](https://happygitwithr.com/) instructions.

## [GitHub](https://happygitwithr.com/github-acct.html)

-   Register a [GitHub](https://github.com/) account.

-   Turn on two-factor authentication under: Settings - Account security.

## [Git](https://happygitwithr.com/install-git.html)

-   Install Git which can be done through [Git for Windows](https://gitforwindows.org).

-   Optionally install a Git client, for example [GitHub Desktop](https://desktop.github.com/), or other [recommended clients](https://happygitwithr.com/git-client.html).

-   In R Studio change terminal to open with Git Bash: Tools - Global Options - Terminal - New terminals open with - Git Bash.

Check Git installed on local machine

```{bash}
where git
git --version
```

## [Set Git details](https://happygitwithr.com/hello-git.html)

**usethis::use_git_config** set Git user name and email

```{r}
usethis::use_git_config(user.name = "USERNAME", user.email = "EMAIL")
```

Check Git user name and email details

```{bash}
git config --global --list
```

## [Connect RStudio to Git](https://happygitwithr.com/rstudio-see-git.html)

-   Tools - Global Options - Git/SVN - Git executable: set path, eg C:/Program Files/Git/bin/git.exe. Ensure **Enable version control interface for RStudio projects** is ticked.

-   Check for [common issues and solutions](https://happygitwithr.com/troubleshooting.html) in Jenny Bryan's instructions.

## [Cache credentials for HTTPS](https://happygitwithr.com/credential-caching.html)

**usethis::create_github_token** Get Personal access taken (PAT), storing token in a password management system

```{r}
usethis::create_github_token()
```

**gitcreds::gitcreds_set** Store credentials in Git credential store

```{r}
gitcreds::gitcreds_set()
gitcreds::gitcreds_get()
```

## [Clone repository from GitHub](https://happygitwithr.com/new-github-first.html)

-   Create a repository under Repository.

-   GitHub: Code - Copy HTTPS URL to clipboard.

-   RStudio: File - New Project - Version Control - Git - Paste HTTPS URL from clipboard.

```{bash}
git clone HTTPS URL
```

## [Clone RStudio project to GitHub](https://happygitwithr.com/existing-github-last.html)

**usethis::use_git** Makes project a Git repository

```{r}
usethis::use_git()
```

-   Commit files to Git repository: RStudio - Git - Tick staged box - Commit.

**usethis::use_github** Create a new repository on GitHub

```{r}
usethis::use_github()
```

-   Pull files from GitHub repository: RStudio Pull button.

-   Push committed files to GitHub: RStudio Push button.

```{bash}
git add -A
git commit -m "COMMIT MESSAGE"
git pull
git push
```


