---
title: "Git notes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Git notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval=FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

These notes were made whilst going through Jenny Bryan's [Happy Git and GitHub for the useR](https://happygitwithr.com/) instructions.

## [GitHub](https://happygitwithr.com/github-acct.html)

-   Register a [GitHub](https://github.com/) account.

-   Turn on two-factor authentication under: Settings - Account security.

-   Create a test repository under Repostitory.

## [Git](https://happygitwithr.com/install-git.html)

-   Install Git which can be done through [Git for Windows](https://gitforwindows.org).

-   Optionally install a Git client, for example [GitHub Desktop](https://desktop.github.com/), or other [recommended clients](https://happygitwithr.com/git-client.html).

-   In R Studio change terminal to open with Git Bash: Tools - Global Options - Terminal - New terminals open with - Git Bash.

Check Git installed on local machine

```{}
where git
git --version
```

## [Set Git details](https://happygitwithr.com/hello-git.html)

**usethis::usethis::use_git_config** set Git user name and email

```{r}
usethis::use_git_config(user.name = "USERNAME", user.email = "EMAIL")
```

Check Git user name and email details

```{}
git config --global --list
```

## [Connect RStudio to Git](https://happygitwithr.com/rstudio-see-git.html)

-   Tools - Global Options - Git/SVN - Git executable: set path, eg C:/Program Files/Git/bin/git.exe. Ensure **Enable version control interface for RStudio projects** is ticked.

-   Check for [common issues and solutions](https://happygitwithr.com/troubleshooting.html) in Jenny Bryan's instructions.

## [Cache credentials for HTTPS](https://happygitwithr.com/credential-caching.html)

**usethis::create_github_token** Get Personal access taken (PAT), storing token in a password management system

```{r}
usethis::create_github_token()
```

**gitcreds::gitcreds_set** Store credentials in Git credential store

```{r}
gitcreds::gitcreds_set()
gitcreds::gitcreds_get()
```

## [Clone repository from GitHub](https://happygitwithr.com/rstudio-git-github.html)
