---
title: "Linear regression notes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear regression notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This example using linear regression to explore the relationship between the body mass and flipper length of palmer penguins, taken from the [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/) package, was created whilst reading

-   Chapter 3 of [An Introduction to Statistical Learning with Applications in R](https://www.statlearning.com/), second edition, by Gareth James, Daniela Witten, Trevor Hastie and Robert Tibshirani

-   [Statistical Inference via Data Science. A ModernDive into R and the Tidyverse](https://moderndive.com/), by Chester Ismay and Albert Kim

```{r echo=FALSE, warning=FALSE, fig.width=4, fig.height=4}
suppressPackageStartupMessages({
  suppressWarnings({
    library(palmerpenguins)
    library(dplyr)
    library(ggplot2)
  })
})

penguins_mass_flipper <- penguins %>%
  select(species, flipper_length_mm, body_mass_g)

penguins_mass_flipper_plot <- ggplot(data = penguins_mass_flipper,
                                     aes(x = flipper_length_mm,
                                         y = body_mass_g)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Penguin size, Palmer Station LTER",
       subtitle = "Flipper length and body mass for 3 species of penguins",
       x = "Flipper length (mm)",
       y = "Body mass (g)") +
  theme(plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")
penguins_mass_flipper_plot
```




 
