---
title: ""
output: html_document
params:
  species_name: NA_character_
  island_name: NA_character_
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo = FALSE}
suppressPackageStartupMessages({
  library(store)
  suppressWarnings({
    library(palmerpenguins)
    library(dplyr)
    library(purrr)
    library(tidyr)
    library(stringr)
    library(gt)
    library(janitor)
  })
})
```

```{r echo = FALSE}
# Assign parameterized values
species_name <- params$species_name
island_name <- params$island_name

# Filter on parameterized values
penguin_data <- penguins %>% 
  filter(species == species_name, island == island_name) %>% 
  select(year, sex, body_mass_g, contains("mm")) %>% 
  drop_na()

# Get years
year_value <- penguin_data %>% 
  distinct(year) %>% 
  pull()
```

## `r species_name` penguins on `r island_name` island

The following table lists the body mass, bill and flipper measurements taken<br/> on
`r nrow(penguin_data)` `r species_name` penguins living on `r island_name` island between `r min(pull(penguin_data, year))` and `r max(pull(penguin_data, year))`.<br/> Penguins with missing measurements are not included.

```{r }
# table
table_measurements <- function(data){
  data %>% 
    gt() %>%
    tab_header(title = md(str_glue("**{species_name} penguins**")),
               subtitle = str_glue("{island_name} island")) %>% 
    cols_label(
      sex = html("Sex<br/><br/>"),
      body_mass_g = html("Body mass<br/>grams"),
      bill_length_mm = html("Bill length<br/>mm"),
      bill_depth_mm = html("Bill depth<br/>mm"),
      flipper_length_mm = html("Flipper length<br/>mm"),
    ) %>% 
    cols_align(
      align = "left",
      columns = everything()
    ) %>% 
    tab_options(
      heading.background.color = "#2C3E50",
      column_labels.font.weight = "bold",
      table.align = "left",
      table.width = pct(50),
      table.font.size = "smaller",
      data_row.padding = px(0)
    )
}
```

```{r echo=FALSE, results='asis'}
# get penguin measurements table for each year
# https://bookdown.org/yihui/rmarkdown-cookbook/child-document.html
child_template <- function(year_value) {
  knitr::knit_child(text = c(
    '#### **Penguin measurements taken during `r year_value`**',
    '',
    '```{r}',
    'penguin_data_year <- penguin_data %>%',
    '   filter(year == year_value) %>%',
    '   select(-year)',
    '',
    'penguin_data_year %>% table_measurements()',
    '```',
    '<br/>',
    ''
  ), envir = environment(), quiet = TRUE)
}
tables <- map(year_value, child_template)
cat(unlist(tables), sep = '\n')
```















