---
title: "Validation checks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation checks}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages({
  suppressWarnings({
    library(assertr)
    library(assertive)
    library(assertthat)
    library(palmerpenguins)
    library(janitor)
    library(data.validator)
  })
})
```

## Validation checks

Data validation is performed using the
[assertr](https://cran.r-project.org/web/packages/assertr/vignettes/assertr.html)
package, with extra checks provided by the
[assertive](https://cran.r-project.org/web/packages/assertive/vignettes/checklists.html)
package. The [assertthat](https://github.com/hadley/assertthat) package
can also be used as an alternative to stopifnot().

All examples perform checks on the [palmerpenguins
dataset](https://allisonhorst.github.io/palmerpenguins/articles/intro.html)
and return TRUE if validation check passes and stops with an error if
the validation check fails.

Validation results can also be added to a validation report, using
appsilon's
[data.validator](https://appsilon.github.io/data.validator/index.html)
package. This report can then be saved and displayed at the end of the
script.

```{r}
# create empty report
report <- data_validation_report()
```

The file path to the validation checks markdown document exists

```{r}
# print to console
assert_that(file.exists("validation_checks.Rmd"), 
            msg = "Path to validation check markdown document is UNSUCCESSFUL")
```

The penguin dataset is not empty

```{r}
# print to console
see_if(not_empty(penguins), 
       msg = "Penguins dataset is EMPTY")
```

The penguins dataset contains 344 rows

```{r}
# print to console
penguins %>%  
  verify(nrow(.) == 344,
         success_fun = success_logical, error_fun = error_stop)
```

The penguins dataset contains eight named columns. This uses a
functional sequence to store verify functions for each column name and
number of columns.

```{r}
# print to console
check_penguins_column_names <- . %>% 
  chain_start() %>% 
  verify(has_all_names("species")) %>%
  verify(has_all_names("island")) %>%
  verify(has_all_names("bill_length_mm")) %>%
  verify(has_all_names("bill_depth_mm")) %>%
  verify(has_all_names("flipper_length_mm")) %>%
  verify(has_all_names("body_mass_g")) %>%
  verify(has_all_names("sex")) %>%
  verify(has_all_names("year")) %>%
  verify(ncol(.) == 8) %>% 
  chain_end(success_fun = success_logical, error_fun = error_stop)
  
penguins %>% 
  check_penguins_column_names
```

Penguins raw dataset date_egg column is a date type

```{r}
# print to console
penguins_raw |>  
  clean_names() |>  
  assert(assertive::is_date, date_egg,
         success_fun = success_logical, error_fun = error_stop)

# display as report
penguins_raw |> 
  clean_names() |> 
  validate(name = "Validation report for raw penguins dataset")  |> 
  validate_cols(is_date, date_egg, 
                description = 'Penguins raw dataset date_egg column is a date type') |> 
  add_results(report)
```

Combination of sample number and species name in penguins raw dataset
are unique

```{r}
# print to console
penguins_raw  |>  
  clean_names() |> 
  assert_rows(col_concat, is_uniq, sample_number, species,
              success_fun = success_logical, error_fun = error_stop)

# display as report
penguins_raw |> 
  clean_names() |> 
  validate(name = "Validation report for raw penguins dataset") |> 
  validate_rows(col_concat, is_uniq, sample_number, species, 
                description = "Combination of sample number and species name in penguins raw dataset are unique") |> 
  add_results(report)
```

All penguins are assigned to a species and an island (they is no missing
data in the species or island columns)

```{r}
# print to console
penguins |> 
  assert(not_na, species, island,  
         success_fun = success_logical, error_fun = error_stop)

# display as report
penguins|> 
  validate(name = "Validation for penguins dataset") |> 
  validate_cols(not_na, species, island, 
                description = "All penguins are assigned to a species and an island (they is no missing data in the species or island columns)") |> 
  add_results(report)
```

All penguins are either male or female (set parameters allow.na = TRUE
to exclude NAs and inverse = TRUE to check if all penguins do not
contain either male or female)

```{r}
# print to console
penguins |>  
  assert(in_set("male", "female"), sex,
         success_fun = success_logical, error_fun = error_stop)

# display as report
penguins|> 
  validate(name = "Validation for penguins dataset") |> 
  validate_cols(in_set("male", "female"),  sex,
                description = "All penguins are either male or female") |> 
  add_results(report)
```

All penguins weight are positive values

```{r}
# print to console
penguins |> 
  dplyr::filter(!is.na(body_mass_g)) %>% 
  assert(assertive::assert_all_are_positive, body_mass_g,
         success_fun = success_logical, error_fun = error_stop)

# display as report
penguins|> 
  validate(name = "Validation for penguins dataset") |> 
  validate_if(body_mass_g > 0,
              description = "All penguins weight are positive values") |>
  add_results(report)
```

All penguins weigh between 2700 and 6300 grams, inclusively

```{r}
# print to console
penguins |> 
  assert(within_bounds(2700, 6300), body_mass_g,
         success_fun = success_logical, error_fun = error_stop)

# display as report as warning
penguins|> 
  validate(name = "Validation for penguins dataset") |> 
  validate_cols(skip_chain_opts = TRUE,
                error_fun = warning_append, within_bounds(2700, 3600), body_mass_g,
                description = "All penguins weigh between 2700 and 3600 grams inclusively") |> 
  add_results(report)
```

## Data exploration

Find the rows in the penguins dataset that are missing all the four
bill, flipper and body mass measurements. Row numbers are displayed in
the index column

```{r}
# print to console
penguins %>% 
  assert_rows(num_row_NAs, function(.x){.x != 4}, bill_length_mm:body_mass_g,
               success_fun = success_continue, error_fun = just_warn)

# display as report
penguins |> 
  validate(name = "Validation for penguins dataset") |> 
  validate_rows(num_row_NAs, function(.x){.x != 4}, bill_length_mm:body_mass_g,
                description = "Each penguin has at least one out of the four bill, flipper and body mass measurements") |> 
  add_results(report)
```

## Validation report

```{r}
print(report)
```

```{r eval = FALSE}
save_report(report)
browseURL("validation_report.html")
```
