---
title: "Validation checks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation checks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages({
  suppressWarnings({
    library(assertr)
    library(assertive)
    library(assertthat)
    library(palmerpenguins)
    library(janitor)
  })
})
```

## Validation checks

Data validation is performed using the [assertr](https://cran.r-project.org/web/packages/assertr/vignettes/assertr.html) 
package, with extra checks provided by the [assertive](https://cran.r-project.org/web/packages/assertive/vignettes/checklists.html)
package. The [assertthat](https://github.com/hadley/assertthat) package can also be used 
as an alternative to stopifnot(). All examples perform checks on the [palmerpenguins dataset](https://allisonhorst.github.io/palmerpenguins/articles/intro.html) and return 
TRUE if validation check passes and stops with an error if the validation check fails

The file path to the palmerpenguins image exists
```{r}
assert_that(file.exists("palmerpenguins_image.png"), 
            msg = "Path palmerpenguins image is UNSUCCESSFUL")
```

The penguin dataset is not empty
```{r}
see_if(not_empty(penguins), 
       msg = "Penguins dataset is EMPTY")
```

The penguins dataset contains 344 rows
```{r}
penguins %>% 
  verify(nrow(.) == 344,
         success_fun = success_logical, error_fun = error_stop)
```

The penguins dataset contains eight named columns
```{r}
penguins %>% 
  chain_start %>%
    verify(has_all_names("species")) %>%
    verify(has_all_names("island")) %>%
    verify(has_all_names("bill_length_mm")) %>%
    verify(has_all_names("bill_depth_mm")) %>%
    verify(has_all_names("flipper_length_mm")) %>%
    verify(has_all_names("body_mass_g")) %>%
    verify(has_all_names("sex")) %>%
    verify(has_all_names("year")) %>%
    verify(ncol(.) == 8) %>%
  chain_end
```

Penguins raw dataset date_egg column is a date type
```{r}
penguins_raw %>% 
  clean_names() %>% 
  assert(assertive::is_date, date_egg,
         success_fun = success_logical, error_fun = error_stop)
```

Combination of sample number and species name in penguins raw dataset are unique
```{r}
penguins_raw %>% 
  clean_names() %>%  
  assert_rows(col_concat, is_uniq, sample_number, species,
              success_fun = success_logical, error_fun = error_stop)
```

All penguins are assigned to a species and an island (they is no missing data in 
the species or island columns)
```{r}
penguins %>% 
  assert(not_na, species, island,  
         success_fun = success_logical, error_fun = error_stop)
```

All penguins are either male or female (set parameters allow.na = TRUE to exclude 
NAs and inverse = TRUE to check if all penguins do not contain either male or female)
```{r}
penguins %>% 
  assert(in_set("male", "female"), sex,
         success_fun = success_logical, error_fun = error_stop)
```

All penguins weight are positive values
```{r}
penguins %>% 
  dplyr::filter(!is.na(body_mass_g)) %>% 
  assert(assertive::assert_all_are_positive, body_mass_g,
         success_fun = success_logical, error_fun = error_stop)
```

All penguins weigh between 2700 and 6300 grams, inclusively
```{r}
penguins %>% 
  assert(within_bounds(2700, 6300), body_mass_g,
         success_fun = success_logical, error_fun = error_stop)
```

## Data exploration

Find the rows in the penguins dataset that are missing all the four bill, flipper
and body mass measurements. Row numbers are displayed in the index column
```{r}
penguins %>% 
  assert_rows(num_row_NAs, function(.x){.x != 4}, bill_length_mm:body_mass_g,
               success_fun = success_continue, error_fun = just_warn)

```




